Evaluating the Performance and Energy Efficiency of the \textsc{COSMO-ART} Model System

### 2014-05-30 paper_enahpc14/RebuttalLetter/rebuttal.tex

It could have been very beneficial to describe the nature of the application in detail instead of 
citing them. I assume that authors have resorted to that technique for the reasons of brevity  - yet 
I beleive that such a detailed description is essential to understand the behaviour of the application.

%%% *** Joseph ***
\begin{weSay}
blabla
\end{weSay}

With missing crucial details of the application and in the abseence of any results towards various 
problem sizes or domains of cases, it is very difficult to do any post-analysis on the overall 
behaviour of the application - for example its runtime scalability (both weak and strong) vs energy 
scalability etc.

%%% *** Will ***
\begin{weSay}
Indeed, there is a huge amount which could be said about COSMO, and COSMO-ART in particular, about strong and weak scalability, energy scalability, along with the underlying science.  However, going into this detail would make this an entirely different paper, and one that has, at least for COSMO, already been published in several forums.   Moreover, such a comprehensive analysis would have taken far more than eight pages.

We had a modest goal in this paper of evaluating the COSMO-ART configuration, which is a benchmark for production at MeteoSwiss, in terms of energy-to-solution and time-to-solution.  To this end we have used state-of-the-art profiling and energy measurement equipment.   In subsequent papers, we plan to address a wider spectrum of energy issues.    
\end{weSay}

@@

In summary, although the paper is very interesting, the lack or depth of information in there severely 
limits the re-usable knowledge contributed by the paper.

%%% *** Will ***
\begin{weSay}
We would be happy to add specific technical details as requested by the reviewers, but the page limit constrains the amout of general detail we can add.  Information about COSMO and COSMO-ART, as well as the tools employed, can be found in the appropriate citations in the paper.  We have chosen to concentrate on documenting the aspects which are actually new to this paper, namely the application of the energy profiling tools to this non-trivial application, and the T2S and E2S evaluation of the COSMO-ART benchmark provided for MeteoSwiss.  
\end{weSay}

@@

\newpage
\section*{Reviewer \#3}

The authors describe the extended atmospheric model COSMO-ART. They present
performance and energy measurements on a selection of HPC systems. There
is no new methodology presented in this paper.

%%% *** Will ***
\begin{weSay}
We do agree that this paper is not intended to develop a methodology for evaluating applications, but rather to use existing tools and methodology to evaluate a complex, real-world application, which brings with it a new spectrum of issues, like additional measurement complexity, large trace files, and the need for system-integrated power measurement tools.
\end{weSay}

@@

Related work is weak. I don't see how [20] and [11] are related to the content
of this paper (other algorithms, other architectures, no comparison of the
energy measurement infrastructure). I suggest to focus more on infrastructure
and tool support for system level power measurement in HPC environments in
the related work section. This would give you the opportunity to argue that
the 1 Hz resolution in one of your test systems is actually state-of-the-art.

%%% *** Will/Joseph ***
\begin{weSay}
Indeed, there is only a tangential relationship between [11] and [20] and our paper, and we would be happy to remove these.  There is in fact very little related work on energy evaluations of real applications; most work in the field limits itself to artificial toy examples, which is definitely not our topic of interest.  We are trying to find more references on infrastructure and tool support, and hope to add these in the final version

 [Jo/Manuel/Sandra: can you find anything appropriate?]
\end{weSay}

@@

In 4.2.2
There is no information about the accuracy

%%% *** Manuel/Sandra ***
\begin{weSay}
We have added information about the accuracy of the PDU.
\end{weSay}

In 4.3
Only -O3 is not appropriate for the Sandy bridge and Ivy Bridge system. At
least -mavx should be added. It looks like some low hanging fruit is reserved
to reach the targeted 5x improvement

%%% *** Will ***
\begin{weSay}
The configuration was taken from the MeteoSwiss benchmark, and was chosen by the developers. We deliberately chose to use their configuration directly to form the baseline, since this is how it would be run in production if running today.  You are right that there is some low-hanging fruit, but optimizations like -mavx alone will not attain the 5x improvement.  
\end{weSay}

@@

It is confusing to compare "Ivy Bridge EP" against "Xeon E5" in the result
discussion, as the Ivy Bridge processor is also called Xeon E5 (v2) as well.

%%% *** Joseph ***
\begin{weSay}
We have fixed this issue along the manuscript.
\end{weSay}

Since the SB and IB measurements use different amounts of cores it is essential to
know if the application scales linearly with the core count. If the application
is scalable 1020 cores at 2.2 GHz should be faster than 672 cores at 2.6 GHz.

%%% *** Will ***
\begin{weSay}
[Jo:  you have to try to answer this: there is information on the scalability of COSMO at least. ]
\end{weSay}

### 2014-05-30 trunk/paper_enahpc14/RebuttalLetter/rebuttal.tex

- It would be easier to compare values in Figure 4 if both energy ETSs and TTSs where together. 
Right now you show ETS TTS ETS TTS, and ETS ETS TTS TTS would be easier to understand.

%%% *** Joseph ***
\begin{weSay}
We have swapped the corresponding columns in the Figure 4, now ETS and TTS appear grouped.
\end{weSay}

- In multiple places: "power dissipation", "power draw". I assumed it this way, if it is not 
I suggest to explain it better.

%%% *** Manuel ***
\begin{weSay}
We have accordingly replaced ``power dissipation'' by ``power consumption'' along the manuscript.
\end{weSay}

%\newpage
\section*{Reviewer \#2}

@@

%%% *** Will ***
\begin{weSay}
Indeed, there is a huge amount which could be said about COSMO, and COSMO-ART in particular, about strong and weak scalability, energy scalability, along with the underlying science.  However, going into this detail would make this an entirely different paper, and one that has, at least for COSMO, already been published in several forums.   Moreover, such a comprehensive analysis would have taken far more than eight pages.

We had a modest goal in this paper of evaluating the COSMO-ART configuration, which is a benchmark for production at MeteoSwiss, in terms of energy-to-solution and time-to-solution.  To this end we have used state-of-the-art profiling and energy measurement equipment.   In subsequent papers, we plan to address a wider spectrum of energy issues.    
\end{weSay}

In summary, although the paper is very interesting, the lack or depth of information in there severely 
limits the re-usable knowledge contributed by the paper.

%%% *** Will ***
\begin{weSay}
We would be happy to add specific technical details as requested by the reviewers, but the page limit constrains the amout of general detail we can add.  Information about COSMO and COSMO-ART, as well as the tools employed, can be found in the appropriate citations in the paper.  We have chosen to concentrate on documenting the aspects which are actually new to this paper, namely the application of the energy profiling tools to this non-trivial application, and the T2S and E2S evaluation of the COSMO-ART benchmark provided for MeteoSwiss.  
\end{weSay}

%\newpage
\section*{Reviewer \#3}

@@

In 4.3
Only -O3 is not appropriate for the Sandy bridge and Ivy Bridge system. At
least -mavx should be added. It looks like some low hanging fruit is reserved
to reach the targeted 5x improvement

%%% *** Will ***
\begin{weSay}
The configuration was taken from the MeteoSwiss benchmark, and was chosen by the developers. We deliberately chose to use their configuration directly to form the baseline, since this is how it would be run in production if running today.  You are right that there is some low-hanging fruit, but optimizations like -mavx alone will not attain the 5x improvement.  
\end{weSay}

In 5.1
The blocking mode does not seem to be blocking. As there are no other
processes/threads available (as many MPI ranks as cores per node) the waiting
ranks will repeatedly call sched\_yield() only to be picked again by the OS
scheduler. This also explains the small benefit as no idle states are used.
You should check the CPU utilization. I guess it is 100\% on all cores in both
modes. If it is not, you should check with powertop which idle states are used.

%%% *** Manuel/Sandra ***
\begin{weSay}
We noticed the problem and fixed it accordingly along the manuscript. We misused blocking/polling terms.
Now we have replaced the ``polling'' and ``blocking''` OpenMPI policies to ``aggressive'' and ``degraded'', 
respectively. The reviewer is completely right on this issue, however even with 
the ``degraded'' policy the CPU utilization is reduced to 50\,\% and thus allows us saving energy.
\end{weSay}

### 2014-05-30 paper_enahpc14/RebuttalLetter/rebuttal.tex

In 4.3
Only -O3 is not appropriate for the Sandy bridge and Ivy Bridge system. At
least -mavx should be added. It looks like some low hanging fruit is reserved
to reach the targeted 5x improvement

%%% *** Will ***
\begin{weSay}
The configuration was taken from the MeteoSwiss benchmark, and was chosen by the developers. We deliberately chose to use their configuration directly to form the baseline, since this is how it would be run in production if running today.  You are right that there is some low-hanging fruit, but optimizations like -mavx alone will not attain the 5x improvement.  
\end{weSay}

In 5.1
The blocking mode does not seem to be blocking. As there are no other
processes/threads available (as many MPI ranks as cores per node) the waiting
ranks will repeatedly call sched\_yield() only to be picked again by the OS
scheduler. This also explains the small benefit as no idle states are used.
You should check the CPU utilization. I guess it is 100\% on all cores in both
modes. If it is not, you should check with powertop which idle states are used.

%%% *** Manuel/Sandra ***
\begin{weSay}
We noticed the problem and fixed it accordingly along the manuscript. We did a misuse of the blocking/polling terms.
Now we have replaced them to ``aggressive'' and ``degraded'', since reflects exactly the experiments of our paper. 
For the second comment, the reviewer is completely right on that issue, however even with  the ``degraded'' policy 
the CPU utilization is reduced to 50\,\% and thus allows us to save energy.
\end{weSay}

### 2014-05-30 trunk/paper_enahpc14/RebuttalLetter/rebuttal.tex

In 5.1
The blocking mode does not seem to be blocking. As there are no other
processes/threads available (as many MPI ranks as cores per node) the waiting
ranks will repeatedly call sched\_yield() only to be picked again by the OS
scheduler. This also explains the small benefit as no idle states are used.
You should check the CPU utilization. I guess it is 100\% on all cores in both
modes. If it is not, you should check with powertop which idle states are used.

%%% *** Manuel/Sandra ***
\begin{weSay}
We noticed the problem and fixed it accordingly along the manuscript. We did a misuse of the blocking/polling terms.
Now we have replaced them to ``aggressive'' and ``degraded'', since reflects exactly the experiments of our paper. 
For the second comment, the reviewer is completely right on that issue, however even with  the ``degraded'' policy 
the CPU utilization is reduced to 50\,\% and thus allows us to save energy.
\end{weSay}

In 5.2
What do you mean by "it aims at computing speed regardless to energy consumption"?
Does it mean it uses Turbo? Please clarify what exactly is different than on the
IB system.

%%% *** Joseph ***
\begin{weSay}
We have fixed the sentence accordingly to avoid any confusion.
\end{weSay}

### 2014-05-30 trunk/paper_enahpc14/RebuttalLetter/rebuttal.tex

%%% *** Manuel/Sandra ***
\begin{weSay}
The authors thank the reviewer for his/her general view of the paper. We consider this work as
a first step towards an energy-efficient version of COSMO-ART inside the Exa2Green project.
\end{weSay}

Some suggestions regarding the paper:

- MONCH and PILATUS are presented as Ivy Bridge and Sandy Bridge clusters respectively in Section 4.1, 
and afterwards in section 5, they are referred as Ivy Bridge and Xeon E5 respectively. Apart from not 
being consistent, both processor families are Xeon E5 thus confusing a bit on what cluster is 
being referred. This also affects Table 1.

%%% *** Joseph ***
\begin{weSay}
We have fixed this issue along the manuscript. To do so we have explained in detail the specifications of the machines
in the Environment setup section and afterwards we refer to the platforms' name or using the
full processor's name plus its microarchitecture.
\end{weSay}

@@

- It would be easier to compare values in Figure 4 if both energy ETSs and TTSs where together. 
Right now you show ETS TTS ETS TTS, and ETS ETS TTS TTS would be easier to understand.

%%% *** Joseph ***
\begin{weSay}
We have swapped the corresponding columns in the Figure 4, now ETS and TTS appear grouped.
\end{weSay}

- In multiple places: "power dissipation", "power draw". I assumed it this way, if it is not 
I suggest to explain it better.

%%% *** Manuel ***
\begin{weSay}
We have accordingly replaced ``power dissipation'' by ``power consumption'' along the manuscript to avoid any confusion.
\end{weSay}

@@

\reviewerSay
The paper investigates the energy footprint of a weather forecasting model on various HPC platforms. 
As such the paper provides the overall energy footprint covering the overall run. In other words the 
granularity is very coarse and is at the application level. Although this will be interesting to the 
COSMO-ART community, the results include some good details on traces  on the
evaluation of polling vs blocking MPI strategies.

%%% *** Manuel/Sandra ***
\begin{weSay}
The authors thank the reviewer for his/her general view of the paper.
\end{weSay}

It could have been very beneficial to describe the nature of the application in detail instead of 
citing them. I assume that authors have resorted to that technique for the reasons of brevity  - yet 
I beleive that such a detailed description is essential to understand the behaviour of the application.

%%% *** Joseph ***
\begin{weSay}
We agree with the rewiever. Indeed we tried to shorten the COSMO-ART's explanation due to space restrictions in the last version of the paper. However
with the page limit extension we added more details about the nature of the application to better understand its behaviour.
\end{weSay}

@@

With missing crucial details of the application and in the abseence of any results towards various 
problem sizes or domains of cases, it is very difficult to do any post-analysis on the overall 
behaviour of the application - for example its runtime scalability (both weak and strong) vs energy 
scalability etc.

%%% *** Will ***
\begin{weSay}
Indeed, there is a huge amount which could be said about COSMO, and COSMO-ART in particular, 
about strong and weak scalability, energy scalability, along with the underlying science.  
However, going into this detail would make this an entirely different paper, and one that has, 
at least for COSMO, already been published in several forums. %   Moreover, such a comprehensive 
%analysis would have taken far more than eight pages.

@@

The authors describe the extended atmospheric model COSMO-ART. They present
performance and energy measurements on a selection of HPC systems. There
is no new methodology presented in this paper.

%%% *** Will ***
\begin{weSay}
We do agree that this paper is not intended to develop a methodology for evaluating applications, but rather to use existing tools and methodology to evaluate a complex, real-world application, which brings with it a new spectrum of issues, like additional measurement complexity, large trace files, and the need for system-integrated power measurement tools.
\end{weSay}

Related work is weak. I don't see how [20] and [11] are related to the content
of this paper (other algorithms, other architectures, no comparison of the
energy measurement infrastructure). I suggest to focus more on infrastructure
and tool support for system level power measurement in HPC environments in
the related work section. This would give you the opportunity to argue that
the 1 Hz resolution in one of your test systems is actually state-of-the-art.

%%% *** Will/Joseph ***
\begin{weSay}
Indeed, there is only a tangential relationship between [11] and [20] and our paper, and we would be happy to remove these.  There is in fact very little related work on energy evaluations of real applications; most work in the field limits itself to artificial toy examples, which is definitely not our topic of interest.  Actually, we have added one reference that helps us to justify the use of the most appropriate wattmeters for an application like COSMO-ART.
%We are trying to find more references on infrastructure and tool support, and hope to add these in the final version
% [Jo/Manuel/Sandra: can you find anything appropriate?]
\end{weSay}

@@

The hardware description is a bit confusing.
- why does the NEC system have 10 racks for 312 nodes if one rack houses 52
     nodes?

%%% *** Joseph ***
\begin{weSay}
We have described in detail the nodes' architectures that this platform is comprised of.
We only used a subset of 52 in our experiments, that is why the reviwer was confused. We now rephrased the explanation in order to be more precise.
\end{weSay}

@@

- what do you mean by "second-generation processor" regarding Sandy Bridge?
     SB is sometimes referred to as 2nd gen Core-i* processor but this does not
     apply to the Xeon line

%%% *** Joseph ***
\begin{weSay}
We recognize the mistake and fix the sentence in the manuscript.
\end{weSay}

- the claim that SB is known to consume more power is wrong. The TDP classes
     are the same for IB and SB (e.g. E5-2670 and E5-2670v2 are both 115W parts)
     It presumably is less efficient, but even such a claim would need a reference

%%% *** Joseph ***
\begin{weSay}
The reviwer is completely right. Thus, we have accordingly modified the text in order to avoid comparisons between SB and IB processors.
\end{weSay}

- It should be added if the Turbo Boost feature is enabled or disabled, this
     seems to be important for the discussion of the TTS results.
- It should be added which power saving mechanisms are used (cpufreq
     governor, available c-states)

%%% *** Manuel/Sandra ***
\begin{weSay}
We have added the details about the Turbo Boost feature and power saving mechanisms for the used platforms.
\end{weSay}

@@

- the second system is described as 16 cores x 42 nodes = 672 cores total. In
     4.3 it is stated that 1344 cores were used (Could be HyperThreading, however
     it is also stated that HT was not considered. Also, 16 MPI ranks per node
     indicate 84 nodes)

%%% *** Joseph ***
\begin{weSay}
We have fixed this mistake. HyperThreading was not considered in our experiments.
\end{weSay}

In 4.2.1
A 5 minute sampling interval seems useless to me. It would be nice to know if
this is actually an instantaneous value every 5 minutes (as I read it) or the
correct average of the past 5 minutes. Figure 5 suggests variation of the power
consumption in the 10\% range. Thus instantaneous values would have a high margin
for error and the 4,6, or 9 samples you use to calculate the average would be
too few to estimate a reliable average.

%%% *** Will/Joseph ***
\begin{weSay}
The E3METER metering framework records averaged power data every 5 minutes, this is as much as we can do in MONCH and PILATUS platforms to collect power consumption data.
On the other hand, Figure 5 represents the power drawn of TINTORRUM, for this case a PDU with a sampling rate of 1 Hz was used instead.
\end{weSay}

@@

In 4.2.2
There is no information about the accuracy

%%% *** Manuel/Sandra ***
\begin{weSay}
We have added information about the accuracy of the APC8653 PDU.
\end{weSay}

@@

In 4.3
Only -O3 is not appropriate for the Sandy bridge and Ivy Bridge system. At
least -mavx should be added. It looks like some low hanging fruit is reserved
to reach the targeted 5x improvement

%%% *** Will ***
\begin{weSay}
The configuration was taken from the MeteoSwiss benchmark, and was chosen by the developers. We deliberately chose to use their configuration directly to form the baseline, since this is how it would be run in production if running today.  You are right that there is some low-hanging fruit, but optimizations like -mavx alone will not attain the 5x improvement.  
\end{weSay}

In 5.1
The blocking mode does not seem to be blocking. As there are no other
processes/threads available (as many MPI ranks as cores per node) the waiting
ranks will repeatedly call sched\_yield() only to be picked again by the OS
scheduler. This also explains the small benefit as no idle states are used.
You should check the CPU utilization. I guess it is 100\% on all cores in both
modes. If it is not, you should check with powertop which idle states are used.

%%% *** Manuel/Sandra ***
\begin{weSay}
We noticed the problem and fixed it accordingly along the manuscript. We did a misuse of the blocking/polling terms.
We have replaced them to ``aggressive'' and ``degraded'', respectively, since exactly reflect the experiments of our paper. 
For the second comment, the reviewer is completely right, however even with  the ``degraded'' policy 
the CPU utilization is still reduced to 50\,\% and thus allows us to save energy.
\end{weSay}

@@

In 5.2
What do you mean by "it aims at computing speed regardless to energy consumption"?
Does it mean it uses Turbo? Please clarify what exactly is different than on the
IB system.

%%% *** Joseph ***
\begin{weSay}
We have rephrased the paragraph accordingly to avoid any confusion.
\end{weSay}

@@

It is confusing to compare "Ivy Bridge EP" against "Xeon E5" in the result
discussion, as the Ivy Bridge processor is also called Xeon E5 (v2) as well.

%%% *** Joseph ***
\begin{weSay}
We have fixed this issue along the manuscript. Now we employ the platform's name for comparing them.
\end{weSay}

Since the SB and IB measurements use different amounts of cores it is essential to
know if the application scales linearly with the core count. If the application
is scalable 1020 cores at 2.2 GHz should be faster than 672 cores at 2.6 GHz.

%%% *** Will ***
\begin{weSay}
We avoided to make comparisons between MONCH and PILATUS platforms - so far we have no more information about the scalability of COSMO-ART. We modified the text accordingly.
%[Jo:  you have to try to answer this: there is information on the scalability of COSMO at least. ]
\end{weSay}

@@

In 5.3
Only 5\% energy reduction when blocked shows that the processors are not idle.
As mentioned earlier, you should add information if power management is
actually used while the processors are waiting.

%%% *** Manuel/Sandra ***
\begin{weSay}
The reviewer is right in this sense. The processors are not idle during blocking MPI routines, however a reduction of the CPU utilization is observed (50\,\%) even with the degraded policy that we used.
\end{weSay}

### 2014-05-30 trunk/paper_enahpc14/Conclusion.tex

We  have   presented  a  methodology  for   comparing  performance  of
COSMO-ART,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandybridge (2.6 GHz) system has
lower  time-to-solution but  higher energy-to-solution  than  an Intel
Ivybridge  (2.2  GHz) system,  although  the  two  metrics are  indeed
strongly  correlated.   A   tool  utilising  Paraver/Extrae  with  new
extensions  for power  consumption  was successfully  applied to  this
non-trivial application.  The  resulting profiles indicate that simple
changes, such as making use of an MPI version based on blocking rather
than polling, can reduce power consumption significantly.

This  reproducible  benchmark provides  a  baseline  for ongoing  work
package within  the EU-funded Exa2Green to  minimise power consumption
of COSMO-ART. Profiling  has given us insight into  the most expensive
code  components,  which are  now  being  altered  to utilise  revised
algorithms.  The  results of these  optimisations will be  reported in
future publications.

>>>>

We  have   presented  a  methodology  for   comparing  performance  of
\cosmoart,  a  regional  weather  forecast  model  augmented  for  the
interactions of  reactive gases and aerosol  particles.  The resulting
benchmarks illustrate  that the  best time-to-solution does  not imply
the best energy-to-solution: an Intel Sandy Bridge (2.6 GHz) system has
lower time-to-solution but higher energy-to-solution than an Intel Ivy
Bridge (2.2 GHz) system, although  the two metrics are indeed strongly
correlated.   On the  other  hand, the  usage  of energy-friendly  MPI
waiting techniques analyzed  with our power-per\-for\-man\-ce framework over
a series of  runs of \cosmoart on \tinto  and power traces demonstrate
that   it  is   possible  to   reduce  both   power   dissipation  and
energy-to-solution   while  maintaining   (or  even   decreasing)  the
time-to-solution. The resulting profiles indicate that simple changes,
such as making use of the  \emph{degraded} MPI policy rather than the \emph{aggressive}
is  possible to  slightly  reduce both  power  dissipation and  energy
consumption (by 2\,\%).

This  reproducible  benchmark provides  a  baseline  for ongoing  work
package within  the EU-funded Exa2Green to  minimize power consumption
of \cosmoart. Profiling  has given us insight into  the most expensive
code  components,  which are  now  being  altered  to utilize  revised
algorithms.  The  results of these  optimizations will be  reported in
future publications.

### 2014-05-30 trunk/paper_enahpc14/Sec4:Experimental-results.tex

Figures~\ref{fig:1} and \ref{fig:2}  account respectively for \monch's
Isola  E1 Rack 2  and \pilat'  Isola HD  total power  measurements for
1-day or 2-day simulations. On \monch, the 1-day simulation was issued
only twice due  to usage restrictions.  As time  resolution was set to
one  update every  5 minutes  for power  sampling, the  averaged power
consumption was computed by considering  6 values for each single run.
On \pilat, the 1-day simulation was  issued four times and a 2-day run
only once.  Similarly, the  averaged power consumption was computed by
considering 4  values for each single  1-day run and 9  values for the
2-days run. Corresponding results are gathered in Table~\ref{tab:1}.

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Monch.eps}
  \caption{Isola E1 Rack 2 total power consumption of \monch.}
  \label{fig:1}
\end{figure}

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.48\textwidth]{Figs/NRJ_benchmark_Pilatus.eps}
  \caption{Isola HD total power consumption of \pilat.}
  \label{fig:2}
\end{figure}

\begin{table}[htbf]
  \centering
  \caption{Averaged power consumption (W) of the platforms.}
  \label{tab:1}
  \begin{tabular}{cccc}
    \hline\noalign{\smallskip}     \textbf{\scriptsize{PILATUS}}     &
    \textbf{\scriptsize{MONCH}}  &  \textbf{\scriptsize{TINTORRUM}}  &
    \textbf{\scriptsize{TINTORRUM}}           \\          &          &
    \textbf{\scriptsize{(\emph{Aggressive})}} & \textbf{\scriptsize{(\emph{Degraded})}}
    \\    \noalign{\smallskip}\hline\noalign{\smallskip}   18035.0   &
    12622.5 & 3713.6 & 3651.8 \\ \noalign{\smallskip}\hline
  \end{tabular}
\end{table}

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As  expected, Intel Sandy Bridge EP processors of \pilat
outperform the Intel Ivy  Bridge EP processors of \monch, being roughly 130\,\%  faster.  The reason
for that is two-fold: \emph{i}) they have higher clock frequency than Ivy
Bridge  (2.6\,GHz  against  2.2\,GHz),   and  \emph{ii})  they  aim  at
computing   speed  regardless   to  be   energy  efficient.    In  our
experiments,  \monch  showed  the  best  energy-to-solution,
reducing the energy consumption of \pilat by approximately 7\,\%.

Averaged TTS and ETS results  over 10 repetitions of 24-hour simulation on
\tinto are shown in  Figure~\ref{fig:4}. The corresponding results for
the  averaged total  power are  also shown  in  Table~\ref{tab:1}.  As
expected, the average total power consumption by \cosmoart while using
the \emph{degraded} MPI policy is about  2\,\% less than that consumed by the
\emph{aggressive} one. This is  due to  the reduction of CPU usage (of about 50\,\%) while using  the \emph{degraded} MPI policy, in
contrast with the \emph{aggressive} policy that always yields to 100\,\% of CPU utilization while waiting for an incoming message. From the point of  view of the execution time, there is a
slight reduction  for the \emph{degraded} MPI  policy. One can  expect that the
costs of the context switches produced in the \emph{degraded} MPI policy can lead to
an  increase  of  the   execution  time,  however,  applications  with
extensive computing,  as e.g.  \cosmoart, may show  better performance
with the \emph{degraded} policy. This effect works in  favor with our results
since the energy consumed is also reduced (2\,\%) just by changing the
behavior of the OpenMPI library  to the \emph{degraded} policy, i.e, with the
\texttt{--mca    mpi\_yield\_when\_idle 1}   parameter   set    in   the
\texttt{mpirun} command.

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    \pilat and \monch for a 24-hour simulation.}
  \label{fig:3}
\end{figure}

\begin{figure}[htbf]
  \centering
  \includegraphics[width=0.5\textwidth]{Figs/Time_E2S_COSMO-ART-1.eps}
  \vspace{-1cm}
  \caption{Mean time-to-solution and  energy-to-solution on \tinto for
    a 24-hour simulation using the  \emph{aggressive} and \emph{degraded} MPI policies. The
    95\,\%  confidence  intervals of  Student's  $t$ distribution  (10
    samples) clearly  indicates an improvement for the \emph{degraded} policy in
    both TTS and ETS.}
  \label{fig:4}
\end{figure}

\begin{figure*}[ht]
  \centering
  \hspace{0.8cm} \scalebox{0.62}{\input{Figs/24hour.pstex_t}}
  \caption{24-hour simulation power  trace  using  the \emph{aggressive}  and
    \emph{degraded} MPI policies on \tinto.}
  \label{fig:5}
\end{figure*}

@@

All  the  traces  in   Figure~\ref{fig:6}  are  divided  in  different
computation  phases. In first  phase all  the processes  perform group
communications      and       execute      MPI-receive      primitives
(e.g. \texttt{MPI\_Recv}). The second one performs computation bursts,
combined  in  between   with  group  communications,  blocking  sends,
immediate  receives and  MPI-wait primitives.  Finally,
group  communications interlaced  with computation  and  a significant
synchronization  phase are  performed. From  the performance  point of
view, one can notice a reduction of 2\,\% in the TTS that \emph{degraded}  MPI
mode leads over the \emph{aggressive} MPI  policy. On the other hand an averaged
reduction of power consumption of  2\,\% is attained with the \emph{degraded} 
MPI policy  only in  routines that actually  perform synchronizations,
group  communications  and  MPI-receive/-wait primitives.   Thus,  the
reduction of the power-energy  pair directly depends on the percentage
of time in which blocking MPI  operations are executed. For the 1 hour
simulation traces, an average of 50\,\% of the time the processes were
waiting for incoming messages.  Taking  into account that  the total power  reduction during
blocking phases  is about 5\,\%,  one can compute an  approximation of
the  power savings,  i.e., $50\,\%  \times  5\,\% =  2.5\%$, which  is
consistent  power  reduction  obtained  for  the  full-day  simulation
(around  2\,\%). However,  we  expect higher  power/energy savings  if
upcoming versions of OpenMPI support full blocking policies as well as if 
future architectures tend proportional computing~\citep{Barroso-2007}.

>>>>

Figures~\ref{fig:1} and \ref{fig:2}  account respectively for \monch's
Isola  E1 Rack 2  and \pilat'  Isola HD  total power  measurements for
1-day or 2-day simulations. On \monch, the 1-day simulation was issued
only twice due  to usage restrictions.  As time  resolution was set to
one averaged update every  5 minutes  for power  sampling, the  averaged power
consumption was computed by considering  6 values for each single run.
On \pilat, the 1-day simulation was  issued four times and a 2-day run
only once.  Similarly, the  averaged power consumption was computed by
considering 4  values for each single  1-day run and 9  values for the
2-days run. Corresponding results are gathered in Table~\ref{tab:1}.

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/NRJ_benchmark_Monch.eps}
  \caption{Isola E1 Rack 2 total power consumption of \monch.}
  \label{fig:1}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/NRJ_benchmark_Pilatus.eps}
  \caption{Isola HD total power consumption of \pilat.}
  \label{fig:2}
\end{figure}

\begin{table}[ht]
  \centering
  \caption{Averaged power consumption (W) of the platforms.}
  \label{tab:1}
  \begin{tabular}{cccc}
    \hline\noalign{\smallskip}     \textbf{\scriptsize{PILATUS}}     &
    \textbf{\scriptsize{MONCH}}  &  \textbf{\scriptsize{TINTORRUM}}  &
    \textbf{\scriptsize{TINTORRUM}}           \\          &          &
    \textbf{\scriptsize{(\emph{Aggressive})}} & \textbf{\scriptsize{(\emph{Degraded})}}
    \\    \noalign{\smallskip}\hline\noalign{\smallskip}   18035.0   &
    12622.5 & 3713.6 & 3651.8 \\ \noalign{\smallskip}\hline
  \end{tabular}
\end{table}

In  Figure~\ref{fig:3}, we  compare both  TTS (right  Y-axis)  and ETS
(left  Y-axis)  metrics  on  both  platforms.  As can be observed, Intel Sandy Bridge EP processors of \pilat
outperform the Intel Ivy  Bridge EP processors of \monch, being roughly 130\,\%  faster. 
% The reason
%for that is two-fold: \emph{i}) they have higher clock frequency than Ivy
%Bridge  (2.6\,GHz  against  2.2\,GHz),   and  \emph{ii}) % they  aim  at
% computing   speed  regardless   to  be   energy  efficient.
In  our experiments,  \monch  showed  the  best  energy-to-solution,
reducing the energy consumption of \pilat by approximately 7\,\%.

Averaged TTS and ETS results  over 10 repetitions of a 1-day simulation on
\tinto are shown in  Figure~\ref{fig:4}. The corresponding results for
the  averaged total  power are  also shown  in  Table~\ref{tab:1}.  As
expected, the average total power consumption by \cosmoart while using
the \emph{degraded} MPI policy is about  2\,\% less than that consumed by the
\emph{aggressive} one. This is  due to  the reduction of the CPU utilization (of about 50\,\%) while using  the \emph{degraded} MPI policy, in
contrast with the \emph{aggressive} one, that always yields to 100\,\% of CPU load while waiting for incoming messages. From the point of  view of the execution time there is a
slight reduction  for the \emph{degraded} MPI  policy. One can  expect that the
costs of the context switches produced in the \emph{degraded} MPI policy can lead to
an  increase  of  the   execution  time,  however,  applications  with
extensive computing,  as e.g.  \cosmoart, may show  better performance
with the \emph{degraded} policy. This effect works in  favor with our results
since the energy consumed is also reduced (2\,\%) just by changing the
behavior of the OpenMPI library  to the \emph{degraded} policy, i.e, with the
\texttt{--mca    mpi\_yield\_when\_idle 1}   parameter   set    in   the
\texttt{mpirun} command.

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/Time_E2S_COSMO-ART-0.eps}
  \caption{Time-to-solution and  energy-to-solution comparison between
    \pilat and \monch for a 1-day simulation.}
  \label{fig:3}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=1\columnwidth]{Figs/Time_E2S_COSMO-ART-1.eps}
  \vspace{-1cm}
  \caption{Mean time-to-solution and  energy-to-solution on \tinto for
    a 1-day simulation using the  \emph{aggressive} and \emph{degraded} MPI policies. The
    95\,\%  confidence  intervals of  Student's  $t$ distribution  (10
    samples) clearly  indicates an improvement for the \emph{degraded} policy in
    both TTS and ETS.}
  \label{fig:4}
\end{figure}

\begin{figure*}[ht]
  \centering
  \hspace{0.8cm} \scalebox{0.62}{\input{Figs/24hour.pstex_t}}
  \caption{1-day simulation power  trace  using  the \emph{aggressive}  and
    \emph{degraded} MPI policies on \tinto.}
  \label{fig:5}
\end{figure*}

@@

All  the  traces  in   Figure~\ref{fig:6}  are  divided  in  different
computation  phases. In first  phase all  the processes  perform group
communications      and       execute      MPI-receive      primitives
(e.g. \texttt{MPI\_Recv}). The second one performs computation bursts,
combined  in  between   with  group  communications,  blocking  sends,
immediate  receives and  MPI-wait primitives.  Finally,
group  communications interlaced  with computation  and  a significant
synchronization  phase are  performed. From  the performance  point of
view, one can notice a reduction of 2\,\% in the TTS that the \emph{degraded}  MPI
policy leads over the \emph{aggressive} one. On the other hand, a
reduction of power consumption of  2\,\% in average is attained with the \emph{degraded} 
policy only for routines that actually  perform synchronizations,
group  communications  and MPI receive and wait primitives.   Thus,  the
reduction of the power-energy  pair directly depends on the percentage
of time in which blocking MPI  operations are executed. For the 1-hour
simulation traces, an average of 50\,\% of the time the processes were
waiting for incoming messages.  Taking  into account that  the total power  reduction during
blocking phases  is about 5\,\%,  one can compute an  approximation of
the  power savings,  i.e., $50\,\%  \times  5\,\% =  2.5\%$, which  is
consistent  power  reduction  obtained  for  the  1-day  simulation
(around  2\,\%). However,  we  expect higher  power/energy savings  if
upcoming versions of OpenMPI support full blocking policies, as well as if 
future architectures tend proportional computing~\citep{Barroso-2007}.
